---
title: Programming with dplyr
date: 2017-10-11
slug: tidyeval
---

If you haven't read the [Programming with
dplyr
vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html)
definitely start there. This post describes how I got comfortable
working with quosures.

dplyr verbs are "referentially transparent" when dealing with the
data.

Take-aways:

1.  When writing dplyr-like verbs use the `.data` pronoun.

The `quo()` function creates quosures.

Use the `unquote()` function to evaluate quosures. `!!`

enquo() uses some dark magic to look at the argument, see what the user typed, and return that value as a quosure. (Technically, this works because function arguments are evaluated lazily, using a special data structure called a promise.)

(If you’re familiar with quote() and substitute() in base R, quo() is equivalent to quote() and enquo() is equivalent to substitute().)


libraries: rlang, tidyeval

Use `!!!` to replace `do.call()`.

When an object keeps track of an environment, it is said to have an enclosure. This is the reason that functions in R are sometimes referred to as closures:

There are three types of unquoting:

1.  basic `UQ()`, `!!` (`UQE()` looks in global environment)
2.  unquote splicing, `UQS()`, `!!!`
3.  unquoting names `:=`

```{r}
library(rlang)

get_help <- function(...) {
    h <- help(...)
    f <- utils:::.getHelpFile(h)
    lines <- capture.output(tools::Rd2txt(f))
    return(lines)
}

lines <- get_help(UQ)
cat(lines[5:28], sep = "\n")
```


```{r}
library(rlang)

UQ
`!!`
UQE
```

I'm not the only one who's confused by this disconnect between the documentation and the code:

https://github.com/tidyverse/rlang/issues/256
